\section{Uniform Continuity}

\begin{callout}
  The first several problems for this unit were multiple choice
  and are omitted here.
\end{callout}

\begin{problem}
  Show directly from the definition that if $f, g$ are uniformly continuous,
  then so is $f+g$.

  \begin{proof}
    Let $\eps > 0$ be given and fix $\eps_{1}, \eps_{2} > 0$ such that
    $\eps_{1} + \eps_{2} = \eps$. By definition, 
    we know:
    \begin{align*}
      \forall \eps_{1} > 0.\, \exists \delta_{1} > 0.\, \forall x, y \in \R.\, \abs{x - y} < \delta_{1} \implies \abs{f(x) - f(y)} < \eps_{1},\\
      \forall \eps_{2} > 0.\, \exists \delta_{2} > 0.\, \forall x, y \in \R.\, \abs{x - y} < \delta_{2} \implies \abs{f(x) - f(y)} < \eps_{2}.
    \end{align*}
    Let $\delta = \min{(\delta_{1}, \delta_{2})}$. Then, with the assistance
    of the triangle inequality, we are able to show that for all $x, y \in R$
    \[
      \abs{x - y} < \delta \implies \abs{(f+g)(x) - (f+g)(y)} \leq \abs{f(x) - f(y)} + \abs{g(x) - g(y)} < \eps_{1} + \eps_{2} = \eps.
    \]
  \end{proof}
\end{problem}

\begin{problem}
  If $f, g$ are uniformly continuous, is $fg$ always uniformly continuous?
  \vspace{\baselineskip}

  No.

  \begin{proof}
    Suppose, to the contrary, that $fg$ is always uniformly continuous. 
    Let $\eps > 0$ be given. Then, put $\delta > 0$ satisfying 
    \[
      \forall x, y \in \R.\, \abs{x - y} < \delta \implies \abs{(fg)(x) - (fg)(y))} < \eps.
    \]
    Notice that
    \[
      \abs{(fg)(x) - (fg)(y)} \leq \abs{g(y)}\abs{f(x)- f(y)} + \abs{f(x)}\abs{g(x)- g(y)}.
    \]
    However, the fixed $\delta$ cannot guarantee that the right side can be made sufficiently small
    (namely, less than $\eps$) for any choice of $x, y \in \R$ when $\abs{x -  y} < \delta$ since
    the coefficients $\abs{g(y)}, \abs{f(x)}$ may not be bounded.
  \end{proof}

  Consider $f(x) = g(x) = x$. It can be shown that $x^{2}$ is not uniformly continuous over $\R$.

\end{problem}

\begin{problem}
  Assume that $g$ is defined on an open interval $(a, c)$ and it is known to be
  uniformly continuous on $(a, b]$ and on $[b, c)$, where $a < b < c$. Prove
  that $g$ is uniformly continuous.

  \begin{proof}
    Let $\eps > 0$ be given and fix $\eps_{1}, \eps_{2} > 0$ such that
    $\eps_{1} + \eps_{2} = \eps$. By definition, 
    we know:
    \begin{align*}
      \forall \eps_{1} > 0.\, \exists \delta_{1} > 0.\, \forall x, y \in (a, b].\, \abs{x - y} < \delta_{1} \implies \abs{g(x) - g(y)} < \eps_{1},\\
      \forall \eps_{2} > 0.\, \exists \delta_{2} > 0.\, \forall x, y \in [b, c).\, \abs{x - y} < \delta_{2} \implies \abs{g(x) - g(y)} < \eps_{2}.
    \end{align*}
    Note that it is sufficient to show that some $\delta > 0$ satisfying the
    conditions above (when substituted for $\delta_{1}$ and $\delta_{2}$)
    also satisfies:
    \[
      \forall x \in (a, b].\, \forall y \in [b, c]).\, \abs{x - y} < \delta \implies \abs{g(x) - g(y)} < \eps_{1} + \eps_{2} = \eps.
    \]

    First, find suitable $\delta_{1}, \delta_{2} > 0$ responses to the
    $\eps_{1}, \eps_{2}$ challenges, respectively. Let $\delta =
    \min{(\delta_{1}, \delta_{2})}$. Then, with the assistance
    of the triangle inequality, we have
    \[
      \abs{g(x) - g(y)} \leq \abs{g(b) - g(x)} + \abs{g(y) - g(b)} < \eps_{1} + \eps_{2} = \eps.
    \]
    whenever $x \in (a, b], y \in [b, c])$ and $\abs{x - y} < \delta$.
  \end{proof}
\end{problem}

